---
title: "PA1_template"
output: html_document

---

#Reproducible Research Project 1

##Loading and preprocessing the data
```{r}
#read in the data
activitydata <- read.csv("activity.csv")
names(activitydata) <- c("steps", "date", "interval")
activitydata$date <- as.Date(activitydata$date)
activitydatanonulls <- activitydata
```

##Q1: What is the mean total number of steps per day?

```{r}
#mean total number of steps per day
#code done
Q1 <- tapply(activitydata$steps, activitydata$date, sum, na.rm=TRUE)
#png("plot1.png", width=480, height=480)
hist(Q1, xlab="Steps", ylab="Frequency", main="Question 1 Graph")
#dev.off()
Q1mean <- mean(Q1, na.rm=TRUE)
Q1median <- median(Q1, na.rm=TRUE)
#paste("The mean for Q1 is ", Q1mean)
#paste("The median for Q1 is ", Q1median)
```

**The median for Q1 is  10395.**
**The mean for Q1 is  9354.22950819672.**

##Q2: What is the average daily activity pattern?

```{r}
#average daily activity pattern
#code done
Q2 <- tapply(activitydata$steps, activitydata$interval, mean, na.rm = TRUE)
#png("plot2.png", width=480, height=480)
plot(as.numeric(names(Q2)), Q2, type = "l", xlab = "5-Minute Intervals", ylab = "Total Steps", main = "Question 2 Graph")
#dev.off()
#paste("Interval with the max number of steps is", max(activitydata$steps, na.rm = TRUE))
```

**Interval with the max number of steps is 806.**

##Q3: Imputing the missing values

```{r}
#imputing missing values
#code done
NAS <- sum(is.na(activitydata$steps))
#paste("There are ", print(NAS), "NA values in the dataset.")
#My strategy for filling in the missing values in the dataset is to populate them with the mean steps value per interval (that are not NAs).
activitydatanonullsmean <- mean(activitydata$steps, na.rm=TRUE)
activitydatanonulls[is.na(activitydatanonulls)]= activitydatanonullsmean
activitydatanonulls <- tapply(activitydatanonulls$steps, activitydatanonulls$date, sum, na.rm=TRUE)
Q3 <- as.numeric(activitydatanonulls)
#png("plot3.png", width=480, height=480)
hist(Q3, xlab="Steps", ylab="Frequency", main="Question 3 Graph")
#dev.off()
Q3meananswer <- mean(Q3, na.rm=TRUE)
Q3median <- median(Q3, na.rm=TRUE)
#paste("The mean for Q3 is ", Q3meananswer)
#paste("The median for Q3 is ", Q3median)
#The mean and median have changed since I replaced the NA values.  The impact of imputing missing data on the estimates of the total daily number of steps have increased.
```

**My strategy for filling in the missing values in the dataset is to populate them with the mean steps value per interval (that are not NAs).**
**There are  2304 NA values in the dataset.**
**The mean for Q3 is  10766.1886792453.**
**The median for Q3 is  10766.1886792453.**

##Q4: Are there differences in activity patterns between weekdays and weekends?

```{r}
#weekdays vs. weekends
for (i in 1:nrow(activitydata)) {
  if (weekdays(activitydata$date[i]) == "Saturday" | weekdays(activitydata$date[i]) == "Sunday") {
    activitydata$isweekday[i] = "yes"
  } else {
    activitydata$isweekday[i] = "no"
  }
}

weekends <- tapply(activitydata[activitydata$isweekday == "no" ,]$steps, activitydata[activitydata$isweekday == "no" ,]$interval, mean, na.rm = TRUE)
weekdays <- tapply(activitydata[activitydata$isweekday == "yes" ,]$steps, activitydata[activitydata$isweekday == "yes" ,]$interval, mean, na.rm = TRUE)
#png("plot4.png", width=480, height=480)
par(mfrow = c(2, 1), mar=c(4,4,0.25,0.25))
#mar=c(bottom side, left side, top side, right side)
plot(as.numeric(names(weekends)), weekends, type = "l", xlab = "5-Minute Intervals Weekends", ylab = "Mean Steps")
plot(as.numeric(names(weekdays)), weekdays, type = "l", xlab = "5-Minute Intervals Weekdays", ylab = "Mean Steps")
#dev.off()
```
